Secure Composition of Robust and Optimising Compilers

### 2023-07-11 popl/main.tex
Proof

\begin{example}[A program not satisfying \gls{tms}]
  Consider the following \texttt{C++11} program that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.

  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
int main() {  // allocates 12 chars containing a greeting message
  char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
  char* to = new char[12];                          // <- address ${\color{gray}\loc_{y}}$

@@

  printf("%cOPL\n", to[6]);
}
  \end{lstlisting}
  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls{tms}, since the location $\loc[_{x}]$ is still allocated.
  However, having returned from \texttt{strncpy}, the \texttt{main} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
\end{example}

@@

\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
\end{theorem}
\MPin{
This you can write the proof of.
Probably even in coq
%
	\MKin{ 
	Should this appear here in the paper? It's basically identical to \Cref{thm:ccab:rtp:ms}

	\MPin{yes it should}}
}

@@

Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
It also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.

>>>>

\begin{example}[A program not satisfying \gls{tms}]
  Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.

  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
int greet() {  // allocates 12 chars containing a greeting message
  char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
  char* to = new char[12];                          // <- address ${\color{gray}\loc_{y}}$

@@

  printf("%cOPL\n", to[6]);
}
  \end{lstlisting}
  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls{tms}, since the location $\loc[_{x}]$ is still allocated.
  However, having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
  It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
\end{example}

@@

\begin{theorem}[Compiler $\ccmsscct$ is secure with respect to \gls{scct}]\label{thm:ccall:rtp:msscct}
  $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$ \Coqed
\end{theorem}
\begin{proof}
  Consequence of \Thmref{thm:ccscct:rtp:scct}, \Thmref{thm:cccfccdce:rtp:ms}, and \Thmref{thm:ccab:rtp:ms} using \Thmref{thm:rtp}.
  \MKin{%
    \textbackslash Thmref or \textbackslash Cref?
  }
% Unfold the definition of $\rtp{\cc{\Ltms}{\Lscct}}{\mssafe\cap\scctsafe}$.
% So, let $\pi\in\mssafe\cap\scctsafe$, $\src{\progvar}$ such that $\rsat{\src{\progvar}}{\pi}$ and let $\obj{\contextvar},\obj{\runtimetermvar}$ with $\progstepto{\prog[\obj]{\obj{\contextvar}}{\left(\cc{\Ltms}{\Lscct}(\src{\progvar})\right)}}{\obj{\runtimetermvar}}{\obj{\trace}}$.
% What is left to show is $\scctfilterLscct[\emptyset]{\obj{\trace}}\in\pi$.
% Let $\irl{\progvar}=\cc{\Ltms}{\Ltrg}\circ\cc{\Ltrg}{\Lms}\circ\cc[\gls{cf}]{\Lms}{\Lms}\circ\cc[\gls{dce}]{\Lms}{\Lms}(\src{\progvar})$ such that $\cc{\Lms}{\Lscct}(\irl{\progvar})=\cc{\Ltms}{\Lscct}(\src{\progvar})$.
% Note that $\pi\in\scctsafe$, so apply \Thmref{thm:ccscct:rtp:scct}
\end{proof}

@@

Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
It also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,necula2005ccured,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.
